Numerical trajectory patterns of floating macroalgae

### 2021-01-11 thesis_chapter_3.Rmd
updating rmd file and output.

In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced. Density distribution maps are an effective method for assessing pathways and ocean connectivity [@Sebille2018]. In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs. The distance that particles traveled was also measured and used to produce boxplots. Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@Sebille2018]. If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@Putman2020; @Putman2018; @Sebille2018].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2:Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3:Comparison of density of particles within each grid cell for the end run time of each scenario. Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas. The bottom plot is the density plot for the passive simulation.', fig.height=10.2, fig.width=7.5}

@@

There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05). The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations amongst simulations.

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4:Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
dis_plot
```

@@

```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5:Comparison of trajectory data between a kelp sphere and kelp cylinder. Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @Sebille2018; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag into account. The passive simulations deviated from passive particles trajectories when including varing levels of hydrodynamic and wind drag scenarios into account. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns and as well our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.

>>>>

In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced. Density distribution maps are an effective method for assessing pathways and ocean connectivity [@sebille2018]. In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs. The distance that particles traveled was also measured and used to produce boxplots. Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@sebille2018]. If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; putman2018; @sebille2018].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag."}
ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density of particles within each grid cell for the end run time of each scenario. Plots A-D repressent the minimum, plots E-H are the mean and plots I-L the maximum cross-sectional areas. The bottom plot is the density plot for the passive simulation.', fig.height=10.2, fig.width=7.5}

@@

There were significant differences among the median distances travelled for all particles in each simulation (Figure 4; p < 0.05). The simulations that considered for 100% hydrodynamic drag travelled less distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations.

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance of all particles from the release site over the course of the simulation for each exposure scenario."}
dis_plot
```

@@

```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 5: Comparison of trajectory data between a kelp sphere and kelp cylinder. Plot A is the comparison of mean trajectories, plot B is the comparison of distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient.", fig.height=8, fig.width=8}

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @sebille2018; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag into account. The passive simulations deviated from passive particles trajectories when including varing levels of hydrodynamic and wind drag scenarios into account. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns and as well our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
